---
title: "Bankruptcy Prediction with Statistical Learning"
subtitle: "STAT432 Final Group Project"
date: "March 21, 2019"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# install and load packages
pkg_list = c('ggplot2', 'tidyr', 'stringr', 'dplyr', 'foreign', 'knitr')
to_install_pkgs = pkg_list[!(pkg_list %in% installed.packages()[,"Package"])]
if(length(to_install_pkgs)) {
  install.packages(to_install_pkgs, repos = "https://cloud.r-project.org")
}
sapply(pkg_list, require, character.only = TRUE)

# Sets default chunk options
knitr::opts_chunk$set(
  fig.align = "center", 
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```


# Group Information

## Group Members

Mary Liu (NetID: zliu203)  
Ziqiao Hua  (NetID: ziqiaoh2)  
Yixin Zhang (NetID: yzhng224)  
Sian Liu (NetID:sianliu2)  
Jiahui Zhao( NetID:jzhao71)  

# Introduction

## Project Description and Problem of Interest
Prediction of firm bankruptcies have been extensively studied in the field of accounting to monitor the financial performance by all shareholders. With the introduction and expansion of statistical learning methods on financial analysis, bankruptcy rate estimation has taken on a new importance [1]. This paper uses an expanded database with more than fifty econometric attributes. The aim of this project is to examine the relationships between these parameters and develop an effective prediction model which allows forecasting the bankruptcy condition of a firm in the near future. In the project, we apply and compare some widely known statistical imputation techniques, such as Decision Tree, K-nearest Neighbor, Logistic Regression and K-Fold Cross Validation and evaluate the performance of these techniques by their accuracy rates. 

## Data Source and Description

The dataset we use is called `Polish Companies Bankruptcy Data Set` which is hosted by UCI Machine Learning Repository and collected from EMIS, a database containing information on emerging markets around the world. The bankrupt companies were analyzed in the period 2000-2012, while the still operating companies were evaluated from 2007 to 2013 [4]. In this project, we will use partial data called `3year` for bankruptcy prediction. It contains financial rates from 3rd year of the forecasting period and corresponding class label that indicates bankruptcy status after 3 years.  

The data `3year` contain contains 64 variables and 10503 observations in total. Below are all the variables which are worth studying. The dependent variable is the class variables with levels 0 or 1, indicating the company bankruptcy or not. Some variables as financial ratio could affect the company be classified as bankruptcy or not. For example, the first variable is "net profit/total assets" which is return on assets (ROA), a financial ratio that shows the percentage of profit a company earns in relation to its overall resources. It is possible that the higher the ROA, the less likely the company will be bankrupt. 

#### Variables

Independent Variables:  
attr1 - net profit / total assets  
attr2 - total liabilities / total assets  
attr3 - working capital / total assets  
attr4 - current assets / short-term liabilities  
attr5 - [(cash + short-term securities + receivables - short-term   liabilities) / (operating expenses - depreciation)] * 365  
attr6 - retained earnings / total assets  
attr7 - EBIT / total assets  
attr8 - book value of equity / total liabilities  
attr9 - sales / total assets  
attr10 - equity / total assets  
attr11 - (gross profit + extraordinary items + financial expenses) / total assets  
attr12 - gross profit / short-term liabilities  
attr13 - (gross profit + depreciation) / sales   
attr14 - (gross profit + interest) / total assets  
attr15 - (total liabilities * 365) / (gross profit + depreciation)   
attr16 - (gross profit + depreciation) / total liabilities   
attr17 - total assets / total liabilities   
attr18 - gross profit / total assets   
attr19 - gross profit / sales   
attr20 - (inventory * 365) / sales   
attr21 - sales (n) / sales (n-1)   
attr22 - profit on operating activities / total assets   
attr23 - net profit / sales   
attr24 - gross profit (in 3 years) / total assets   
attr25 - (equity - share capital) / total assets   
attr26 - (net profit + depreciation) / total liabilities   
attr27 - profit on operating activities / financial expenses   
attr28 - working capital / fixed assets   
attr29 - logarithm of total assets   
attr30 - (total liabilities - cash) / sales   
attr31 - (gross profit + interest) / sales   
attr32 - (current liabilities * 365) / cost of products sold   
attr33 - operating expenses / short-term liabilities   
attr34 - operating expenses / total liabilities   
attr35 - profit on sales / total assets   
attr36 - total sales / total assets   
attr37 - (current assets - inventories) / long-term liabilities   
attr38 - constant capital / total assets   
attr39 - profit on sales / sales   
attr40 - (current assets - inventory - receivables) / short-term liabilities   
attr41 - total liabilities / ((profit on operating activities + depreciation) * (12/365))   
attr42 - profit on operating activities / sales   
attr43 - rotation receivables + inventory turnover in days   
attr44 - (receivables * 365) / sales   
attr45 - net profit / inventory   
attr46 - (current assets - inventory) / short-term liabilities   
attr47 - (inventory * 365) / cost of products sold   
attr48 - EBITDA (profit on operating activities - depreciation) / total assets   
attr49 - EBITDA (profit on operating activities - depreciation) / sales
attr50 - current assets / total liabilities   
attr51 - short-term liabilities / total assets   
attr52 - (short-term liabilities * 365) / cost of products sold)
attr53 - equity / fixed assets   
attr54 - constant capital / fixed assets   
attr55 - working capital   
attr56 - (sales - cost of products sold) / sales   
attr57 - (current assets - inventory - short-term liabilities) / (sales - gross profit - depreciation)   
attr58 - total costs /total sales   
attr59 - long-term liabilities / equity   
attr60 - sales / inventory   
attr61 - sales / receivables   
attr62 - (short-term liabilities *365) / sales    
attr63 - sales / short-term liabilities    
attr64 - sales / fixed assets   

Dependent Variable:  
class - the response variable Y: 0 = did not bankrupt; 1 = bankrupt  

#### 10 Observations 

The head ten observations are listed below:  

```{r}
# load in data
year3 = foreign::read.arff('3year.arff')

# top 10 observations
knitr::kable(head(year3,10)[, 1:8], format = 'latex')
knitr::kable(head(year3,10)[, 9:16], format = 'latex')
knitr::kable(head(year3,10)[, 17:24], format = 'latex')
knitr::kable(head(year3,10)[, 25:32], format = 'latex')
knitr::kable(head(year3,10)[, 33:40], format = 'latex')
knitr::kable(head(year3,10)[, 41:48], format = 'latex')
knitr::kable(head(year3,10)[, 49:56], format = 'latex')
knitr::kable(head(year3,10)[, 57:65], format = 'latex')
``` 

# Methods
In this project, various methods which based on statistical hypothesis testing, statistical modeling and statistical learning techniques will be explored. We will start with logistic regression since the dependent variable is binary with only two classical levels. Then, we will apply some of the widely used classification models: decision trees, random forest, K-Nearest Neighbors and Bagging. By applying K-Fold Cross Validation, we will be able to analyze with imputed and resampled datasets. Finally, we will evaluate and rank the performance of models we used on the validation datasets by accuracy (error) rates (with confusion table). We will be able to find the model with best performance on predicting future bankruptcy condition for a firm by the end of the project. Other analysis methods which we have learned from class (e.g. Ridge Regression, Lasso Regression and Hierarchy Clustering) may be also applied in this project. 

# Challenges

We have discovered some challenges. The first challenge is value Missing: by looking through the data, we found missing values for some observations, indicating data imputation techniques like Expectation-Maximization or KNN are needed. The second challenge is data imbalance: the summary of dependent variable suggests among the 10503 observations, only 459 of them reported bankruptcy and the rest 10008 firms did not bankrupt in the future 3 years. This ratio indicates a need to oversample the minority categorical class. The next challenge is risk of overfitting: the data contains more than 60 predictor variables, indicates there is a great risk/potential of overfitting models and tuning parameter is needed in decision tree. Finally, since the data set contains 10503 observations in total, there is a challenge of extensive data visualization and the computational burden may increase. 

# Appendix

## Code Chunks

If code chunks are used within the document, this information can be dynamically retrieved and embedded.

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```

# Reference

[1] Sudheer Chava and Robert A. Jarrow, Bankruptcy Prediction with Industry Effects, Review of Finance 8: 537-569, 2004, http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.495.4409&rep=rep1&type=pdf   

[2] Risck L. Wilson, Ramesh Sharda, Bankruptcy Prediction Using Neural Networks, ScienceDirect, Volume 11, Issue 5, June 1994, Pages 545-557, https://doi.org/10.1016/0167-9236(94)90024-8   

[3] Sai Surya Teja Maddikonda and Sree Keerthi Matta, Bankruptcy Prediction: Mining the Polish Bankruptcy Data, https://github.com/smaddikonda/Bankruptcy-Prediction/blob/master/Bankruptcy%20Prediction%20Report.pdf 

[4] Polish Companies Bankruptcy Data Set, UCI Machine Learning Repository,http://archive.ics.uci.edu/ml/datasets/Polish+companies+bankruptcy+data
